### 2022-04-25 paper/cover-letter/cover-letter.Rmd
Add all of reviewers' comments in response file.

---
from: Dewi Amaliah
fromqualifications: Master of Business Analytics
fromWWW: "[https://numbats.github.io/yowie/](numbats.github.io/yowie/)"
fromemail: dlamaleeah@gmail.com
fromphone: "+61 435 444 025"
fromsignature: sigfile.png
toname: Prof. Nick Horton
toaddress: Editor, JSDSE 
opening: Dear Prof. Horton
closing: Regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider the revised version of our paper titled "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth" for the JDSDE Special Issue on "Teaching Reproducibility and Responsible Workflow".

The paper is co-authored by Prof. Di Cook, Dr Emi Tanaka, Kate Hyde and Dr Nick Tierney. We believe it is a good fit for the special issue because it covers various aspects of data management and curation component of data acumen, including data preparation and cleaning, missing and conflicting data. The code to reproduce the data cleaning, and make further updates of the resulting data is provided.

We also provide the response of editor and reviewers' comments. We also agree if you share this paper with the discussant for this special issue.